Manuscript

### 2019-01-22 manuscript_analysis.Rmd
identified data duplications due to manual MAF annotations, need to fix

Merged mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./data_munging/18Q4/depmap_18Q4_mutation_calls_20181114.csv.gz", header = TRUE, sep = ",", row.names = 1)
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"
# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Wildtype", "Mutant")
maf_df <- unique(maf_df)
saveRDS(maf_df, "./data_munging/rds/maf_df_18Q4.rds", compress = "xz")
```

@@

@@

Pearson correlations for assessing effect of gene expression and copy number on gene essentiality score and AUC.

@@

ggsave("./plots/manuscript/crispr_venn.pdf", crispr_venn)
```

@@

Spearman correlations for gene expression and copy number:
```{r fig.height=6, fig.width=12}
crispr_spearman_res_GE <- transform(crispr_spearman_res_GE, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_spearman_res_GE$p_log10 <- -log10(crispr_spearman_res_GE$p)
crispr_spearman_res_CN <- transform(crispr_spearman_res_CN, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_spearman_res_CN$p_log10 <- -log10(crispr_spearman_res_CN$p)

crispr_ge_spearman_corr_plot <- ggplot(data = crispr_spearman_res_GE, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "darkorchid4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)", subtitle = "CRISPR: Correlation between AUC z-score and gene expression")
crispr_cn_spearman_corr_plot <- ggplot(data = crispr_spearman_res_CN, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "darkorchid4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.title.y = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)", subtitle = "CRISPR: Correlation between AUC z-score and copy number")

@@

drug_venn_mut <- venn.diagram(list("CCLE" = ccle_mut_dgs, "CTRP" = ctrp_mut_dgs, "GDSC" = gdsc_mut_dgs), resolution = 5000, main = "A", main.pos = c(0, 1), main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 50, 120), cat.dist = c(0.001, 0.02, 0.0001), cat.cex = 1.5, cex = 1, filename = NULL, height = 4000)
# in GDSC only
drug_venn_mut[[7]]$label  <- paste(sort(setdiff(setdiff(gdsc_dgs, ctrp_dgs), ccle_dgs)), collapse = "\n")
# in GDSC and CTRP only
drug_venn_mut[[8]]$label <- paste(intersect(gdsc_dgs, ctrp_dgs), collapse = "\n")
# in CTRP only
drug_venn_mut[[9]]$label <- paste(sort(setdiff(setdiff(ctrp_dgs, gdsc_dgs), ccle_dgs)), collapse = "\n")
# in CTRP and CCLE only
drug_venn_mut[[10]]$label <- intersect(ccle_dgs, ctrp_dgs)

@@

# Pancancer vs. Lineage

>>>>

Mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./data_munging/18Q4/depmap_18Q4_mutation_calls_20181114.csv.gz", header = TRUE, sep = ",", row.names = 1)
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"
# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Wildtype", "Mutant")
maf_df <- unique(maf_df)
saveRDS(maf_df, "./data_munging/rds/maf_df_18Q4.rds", compress = "xz")
```

@@

457/3462 gene-cell line combinations have more than one mutation in the MAF file. **FIX THIS**
```{r}
# test <- maf_df_g2p %>% count(Hugo_Symbol, CCLE_Name)
# test <- test[order(test$n, decreasing = TRUE),]
```

@@

Pearson and Spearman correlations for assessing effect of gene expression and copy number on gene essentiality score and AUC.

Example of how to describe Spearman correaltion results: ""A Spearman's correlation was run to determine the relationship between 23 groundwater uranium and TDS values. There was a strong, positive monotonic correlation between Uranium and TDS ( = .71, n = 23, p < .001)."

@@

# ggsave("./plots/manuscript/crispr_venn.pdf", crispr_venn)
```

@@

Spearman correlations for gene expression and copy number:
```{r fig.height=6, fig.width=12}
crispr_spearman_res_GE <- transform(crispr_spearman_res_GE, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_spearman_res_GE$p_log10 <- -log10(crispr_spearman_res_GE$p)
crispr_spearman_res_CN <- transform(crispr_spearman_res_CN, Hugo_Symbol = reorder(Hugo_Symbol, corr)) 
crispr_spearman_res_CN$p_log10 <- -log10(crispr_spearman_res_CN$p)
spearman_KRAS <- rbind(filter(crispr_data_g2p, Hugo_Symbol == "KRAS") %>% summarize(p = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), filter(crispr_data_g2p, Hugo_Symbol == "KRAS") %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"))

crispr_ge_spearman_corr_inset <- ggplot(filter(crispr_data_g2p, Hugo_Symbol == "KRAS"), aes(x = RPKM, y = Score)) +
  geom_point(pch = 21, fill = "cadetblue3", alpha = 0.6) +
  labs(x = "KRAS Gene Expression (RPKM)", y = "CERES Score", subtitle = paste0("Spearman correlation coefficient = ", round(spearman_KRAS$corr[1], 2), " (", printp(spearman_KRAS$p[1]), ")"))

crispr_cn_spearman_corr_inset <- ggplot(filter(crispr_data_g2p, Hugo_Symbol == "KRAS"), aes(x = CN, y = Score)) +
  geom_point(pch = 21, fill = "palegreen2", alpha = 0.6) +
  labs(x = "KRAS Copy Number (log2 ratio)", y = "CERES Score", subtitle = paste0("Spearman correlation coefficient = ", round(spearman_KRAS$corr[2], 2), " (", printp(spearman_KRAS$p[2]), ")"))

crispr_ge_spearman_corr_plot <- ggplot(data = crispr_spearman_res_GE, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "deepskyblue4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)") +
  annotation_custom(ggplotGrob(crispr_ge_spearman_corr_inset), xmin = 12, xmax = 32, ymin = -0.15, ymax = -0.45) +
  geom_curve(aes(x = 3.5, y = -0.285, xend = 13, yend = -0.16), color = "lightgray", curvature = -0.2, arrow = arrow(length = unit(0.03, "npc")))

crispr_cn_spearman_corr_plot <- ggplot(data = crispr_spearman_res_CN, aes(x = Hugo_Symbol, y = corr)) +
  geom_point(aes(fill = p_log10), pch = 21, size = 5) +
  scale_fill_gradient2(low = "white", mid = "white", high = "green4", midpoint = 1.3010299957) +
  coord_cartesian(ylim = c(-0.45, 0.2)) +
  theme(legend.position = "bottom", axis.title.y = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(fill = "-log10(p)", x = "Gene", y = "Spearman Correlation Coefficient (r_s)") +
  annotation_custom(ggplotGrob(crispr_cn_spearman_corr_inset), xmin = 12, xmax = 32, ymin = -0.15, ymax = -0.45) +
  geom_curve(aes(x = 2.5, y = -0.232, xend = 13, yend = -0.16), color = "lightgray", curvature = -0.2, arrow = arrow(length = unit(0.03, "npc")))

@@

drug_venn_mut <- venn.diagram(list("CCLE" = ccle_mut_dgs, "CTRP" = ctrp_mut_dgs, "GDSC" = gdsc_mut_dgs), resolution = 5000, main = "A", main.pos = c(0, 1), main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 50, 120), cat.dist = c(0.001, 0.02, 0.0001), cat.cex = 1.5, cex = 1, filename = NULL, height = 4000)
# in GDSC only
drug_venn_mut[[7]]$label  <- paste(sort(setdiff(setdiff(gdsc_mut_dgs, ctrp_mut_dgs), ccle_mut_dgs)), collapse = "\n")
# in GDSC and CTRP only
drug_venn_mut[[8]]$label <- paste(intersect(gdsc_mut_dgs, ctrp_mut_dgs), collapse = "\n")
# in CTRP only
drug_venn_mut[[9]]$label <- paste(sort(setdiff(setdiff(ctrp_mut_dgs, gdsc_mut_dgs), ccle_mut_dgs)), collapse = "\n")
# in CTRP and CCLE only
drug_venn_mut[[10]]$label <- intersect(ccle_mut_dgs, ctrp_mut_dgs)

@@

# Comopare pancancer vs. lineage
